Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors

### 2019-02-07 paper/ms.tex
Work on intro in Carlyn's paper.

\begin{abstract}
Recent studies of stellar evolution predict the existence of "hybrid" white 
dwarfs, made of a C/O/Ne
core within a O/Ne shell, and that these are viable progenitors for supernovae.
More recent work found that the C/O core is mixed with the surrounding O/Ne
while the white dwarf cools. Inspired by this scenario, we performed simulations of
thermonuclear supernovae from these hybrid progenitors. The progenitor model 
was constructed with the one-dimensional stellar evolution code MESA and
represented a star evolved through the phase of unstable interior mixing 
followed by accretion until it reached conditions for the ignition of 
carbon burning. The MESA model was then mapped to a two-dimensional initial 
condition and an explosion simulated with FLASH. For comparison, a similar
simulation of an explosion was performed from a traditional C/O progenitor
white dwarf. By comparing the yields of the explosions, we found....
how explosions from these models differ from explosions from previous models
without the mixing during the white dwarf cooling.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Thermonuclear (Type Ia) supernovae (\SNeIa) are bright stellar explosions 
thought to occur when approximately 1.0 \Msun\ of material composed principally 
of C and O burns under degenerate conditions. This class of supernovae is
know to synthesize much of the Fe-group elements found in the galaxy and
the light curves of these events have a spectial property that allows
their use as distance indicators for cosmological studies~\citep{phillips:absolute}
resulting in the discovery of the acceleration of the expansion of
the Universe and thus the inference of Dark 
Energy~\citep{riess.filippenko.ea:observational,
perlmutter.aldering.ea:measurements,leibundgut2001}. 
This special property of the light curve is thought to follow
from the fact that the source of luminosity, the radioactive decay
of \Ni{56}, synthesized by the thermonuclear burning, is also the
principle source of opacity~\citep{Pinto2001The-type-Ia-sup}.

Supernovae are classified observationally 
by their light curves and spectra~\citep{minkowski41,bertola64,porterfilippenko87,
wheelerharkness1990conf,Fili97}, with the type Ia designation following from
the absence of H in the spectrum and the presence of a specific Si 
line~\citep{filippenko:optical,hillebrandt.niemeyer:type}. These events
have been associated with C burning under degenerate conditions
for some time~\citep{hoylefowler60,arnett.truran.ea:nucleosynthesis},
but discerning the setting(s) of these events is proving difficult
and remains the subject of active research. At present there are three
widely-accpeted scenarios: the {\em single degenerate} scenario,
the {\em double detonation} or {\em sub-Chandrasekhar} scenario, 
the {\em white dwarf merger} or {\em double degenerate} scenario.
We briefly describe these in the subsection that follows. `
Also see~\citet{hillebrandt.niemeyer:type,howell2011,hillebrandtetal2013,calderetal2013,roepkesim2018}
for additional discussion.

@@

Within this progenitor setting, many mechanisms for the explosion have 
been studied with the result being that the outcome is sensitive to
the ignition conditions and the nature of the thermonuclear buring.
Studies indicate that neither a pure deflagration (subsonic flame front)
nor a pure detonation (supersonic burning front) produce ejecta conisitent
with observations \citep{arnett69,roepkeetal07}.
Instead, the models that best reproduce the observed
stratified ejecta are those in which the burning begins as a 
subsonic deflagration, which allows the star to react and expand
thus lowering the density, that is followed by a 
detonation~\citep{Nomo84,Khokhlov1991Delayed-detonat,HoefKhok96,GameKhokOran05}.
The ``classic'' delayed detonation model is that of 
Khokhlov \cite{Khokhlov1991Delayed-detonat,hoflich.khokhlov.ea:delayed,GameKhokOran05},
and many variations have been 
explored \cite[and references therein]{hillebrandtetal2013,calderetal2013}.
The approach we employ for this study is a variation, the 
deflagration-to-detonation transition (DDT), in which the transition
occurs at the top of a rising Rayleigh-Taylor unstable plume of hot, burned 
material when it reaches a threshold density.
Which we describe this approach and our methodology  in detail below.

The double detionation picture is a variation of the single degenerate picture.
In this case, a white dwarf accretes material from a companion, but rather
than gaining enough mass to approach the Chandrasekhar limit, a detonation
occurs in the accreted layer that subsequently triggers a detonation
in the underlying white dwarf \citet{woosleyweavertaam80,taam80a,taam80b,
nomoto80,nomoto82b}. Early studies indicated that a detonation in an 
accreted He
layer could produce an inwardly propagating
shock that would ignite a detonation in the C-O core and found
that this scenario could work for a wide range of white dwarf
masses, not just the near-Chandrasekhar case \citep{livne90}, hence the
moniker ``sub-Chandrasekhar'' \citep{ww94}. Multidimensional supported the
efficacy of this picture~\citep{livneglasner91, livnearnett95,HoefKhok96,
hoeflichetal96, wigginsfalle97,wigginsetal98,garciasenzbravowoosley99} 
but indicated that uncertainties 
like the initial conditions were found to play a significant role.
on the explosion outcome.

A particular problem was that most models 
included a massive He layer, which leads to 
the unobserved result intermediate and heavy elements synthesized in 
the He detonation in the outer parts of the ejecta~\citep{HoefKhok96, 
hoeflichetal96,finkhillebrandtroepke2007,simetal2010}.
\citet{bildstenetal2007}, however, found that fairly thin He layers
could flash on sub-Chandrasekhar
mass white dwarfs, partially resolving the issue and encouraging
further research~\citep{simetal2012,brooksetal2015, shenetal2018,
glasneretal2018}.

The white dwarf merger progentor picture has two white dwarfs coming 
together and subsequently exploding~\citep{tutukovyungelson76,tutukovyungelson79,
webbink84,ibentutukov84}. This scenario provides an abundance of degenerate fuel, 
which may explain some bright events~\citep{scalzo:2010,Yuan:2010}. 
An early concern about this model came from studies finding that
the white dwarf will ignite at the edge and the coalesced object, 
and a flame will burn inward, converting the C-O white dwarf into 
an O-Ne-Mg white dwarf~\citep{saionomoto1985,saionomoto2004}, with
further accretion leading to the collapse of the white dwarf
into a neutron star, a scenario termed ``accretion-induced collapse''
\citep{nomotokondo1991}.

Subsequent work allayed this concern~\citep{yoonetal2007,lorenaguilaretal2009,
Shenetal12, pakmoretal2012b} and motivated by the need to expain briht events
and the estimated population WD pairs, active research continues. Contemporary
research focuses on variations on the merger idea, including 
inspiraling pairs, collisions, violent mergers,
and the ``core-degenerate'' model in which the merger takes place in
a common envelope \citep{raskinetal2009,pakmoretal2011,kashi:2011,pakmoretal2012a,Shenetal12,katzetal2016}.

@@

Our models are within the DDT explosion paradigm \cite{1986SvAL,
Khokhlov1991Delayed-detonat,NiemWoos97,Niem99,belletal2004,fishjump2015}. 
In this scenario, the accretion of mass on the white dwarf compresses and 
heats the core, igniting carbon fusion and driving a period of convection
\citep{WoosWunsKuhl04,wunschwoosley2004,Kuhletal06,nonakaetal2012}.
At some point, the fusion rate becomes fast enough due to the rising
temperature that energy production exceeds convective cooling and
the deflagration phase begins in the core~\citep{Nomo84,WoosWunsKuhl04}.
This flame is unstable, and as the deflagration propagates toward the surface
of the WD, it is subject to the Rayleigh-Taylor instability that generates
turbulence and boosts buring.  
Burning proceeds as a deflagration for about one second, and then 
the flame transitions to a detonation~\citep{hoflich.khokhlov.ea:delayed}.

@@

The mechanism by which the deflagration transistions to a detonation
is imcompletely understood, but the development of fluid instabilities
at the deflagration front, particuarly the Rayleigh-Taylor 
instability~\cite{taylor+50,chandra+81}, is an accepted part of the process.
Many ideas have been proposed, including the deflagration front
entering a regime of distributed burning and the net
rate becoming supersonic \cite{NiemWoos97}, the unstable
flame reaching a certian fractal dimension~\cite{woosley90},
and the Zel'dovich mechanism in which condtions are right for
a pocket of fuel to 
detonate~\cite{zeldovichetal1970,KhokOranWhee97,jacketal2014}.
Our simulations assume that the transition occurs when the top of a 
rising, Rayleigh-Taylor unstable plume reaches a characteristic low
density \cite{townsley.calder.ea:flame}.

@@

Modern computing resources now enable simulations with unprecedented 
realism, allowing both one-dimensional simulations with a vast
amount of included physics and full three-dimensional simulations
albeit with less included physics~\cite{caldertownsley2018}.
In the area of stellar evolution, recent investigations revisiting 
late-time evolution of roughly 8 \Msun stars indicate that under
the right circumstances, ``hybrid" white dwarfs having a C/O core surrounded by O/Ne 
mantle may form \citep{siess2009,denissenkovetal2013}. These hybrid
white dwarf are thought to form when mixing at the lower convective boundary quenches
C burning in an asymptotic giant branch (AGB) star, leaving unburned C
in the core. The situation is at best uncertain, however, and the 
results depending on assumptions about convective
overshoot that have been questioned \citep{chenetal2014,lecoanetetal16,lattanzioetal2017}.

Assuming a hybrid WD forms, there is subsequent evolution that will occur. There
are two ways in which a hybrid WD can become a progenitor of a thermonuclear supernova.
First, the hybrid could be part of a binary system in which the companion star becomes
another white dwarf and the two merge. Recent population synthesis work supports
this pathway by indicating a substantial contribution to the Ia population rate from 
mergers where one member is a hybrid~\citep{yungelsonkuranov2017}. The second way
occurs if the hybrid WD is part of a binary with a main sequence or giant companion
and gains enough mass to approach the Chandrasekhar mass and ignite C fusion (i.e.
the single degenerate picture)~\citep{}.

In either case, the hybrid WD will experience a period of cooling, which will
make the core-mantle interface subject to convective 
instability \citep{brooksetal2017,schwabgaraud2018}. In the case of the 
hybrid WD accreting and approaching the Chandrasekhar mass, accretion will
heat the core and start C fusion, leading to a period of ``simmering" prior
to the explosion \citep{who?}. The upshot is that is likely to be considerable
mixing after the hybrid forms that may homegenize the compostion 
\citep{denissenkovetal2015,brooksetal2017,schwabgaraud2018}.

@@

If we refer to the work by name, it should appear ``The hybrid models
of~\citet{brooksetal2017}." Or, we can just cite it~\citep{brooksetal2017}.

@@

Summarize what we do a la:
In our simulations, we initialize a detonation once the 
deflagration front reaches a
characteristic DDT fuel density, which controls the degree of
expansion the star undergoes during the deflagration stage.

>>>>

\begin{abstract}
The existence of "hybrid" white dwarfs, made of a C/O
core within a O/Ne shell is accepted, and studies indicate that these are 
viable progenitors for thermonuclear (type Ia) supernovae.
Recent work found that the C/O core is mixed with the surrounding O/Ne
while the white dwarf cools. To further investigate the efficacy
of these white dwarfs as progenitors of thermonuclear supernovae, we performed 
simulations of thermonuclear supernovae from a new series of hybrid progenitors
that include the effects of cooling. The progenitor white dwarf model
was constructed with the one-dimensional stellar evolution code MESA and
represented a star evolved through the phase of unstable interior mixing 
followed by accretion until it reached conditions for the ignition of 
carbon burning. This MESA model was then mapped to a two-dimensional initial 
condition and an explosion simulated with FLASH. For comparison, a similar
simulation of an explosion was performed from a traditional C/O progenitor
white dwarf. By comparing the yields of the explosions, we found that as
with earlier studies, these hybrid models are viable supernova progenitors.
We contrast explosions from these models with explosions 
from previous models {\color{red} without the mixing during the white dwarf cooling.}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Thermonuclear (Type Ia) supernovae (\SNeIa) are bright stellar explosions 
thought to occur when approximately 1.0 \Msun\ of material composed principally 
of C and O burns under degenerate conditions. This class of supernovae is
know to synthesize much of the Fe-group elements found in the galaxy, and
the light curves of these events have a special property that allows
standardization of their light curves and thus use
as distance indicators for cosmological studies~\citep{phillips:absolute}.
This use resulted in the discovery of the acceleration of the expansion of
the Universe and thus the inference of Dark 
Energy~\citep{riess.filippenko.ea:observational,
perlmutter.aldering.ea:measurements,leibundgut2001}, and these events
remain the premier distance indicators for cosmological studies {\color{red} REF?}. 
The special property of the light curve is thought to follow
from the fact that the source of luminosity, the radioactive decay
of \Ni{56}, synthesized by the thermonuclear burning, is also the
principle source of opacity~\citep{Pinto2001The-type-Ia-sup}, giving
the Phillips relation between the brightness of an event and the
rate of decline of the B-band magnitude from maximum~\citep{phillips:absolute}.

Supernovae are classified observationally 
by their light curves and spectra~\citep{minkowski41,bertola64,porterfilippenko87,
wheelerharkness1990conf,Fili97}, with the type Ia designation following from
the absence of H in the spectrum and the presence of a specific Si 
line~\citep{filippenko:optical,hillebrandt.niemeyer:type}. These events
have been associated with C burning under degenerate conditions
for some time~\citep{hoylefowler60,arnett.truran.ea:nucleosynthesis},
but discerning the setting(s) of these events is proving difficult
and remains the subject of active research. At present there are three
widely-accpeted scenarios: the {\em single degenerate} scenario,
the {\em double detonation} or {\em sub-Chandrasekhar} scenario, 
the {\em white dwarf merger} or {\em double degenerate} scenario.
We briefly describe these in the subsection that follows.
Also see~\citet{hillebrandt.niemeyer:type,howell2011,hillebrandtetal2013,calderetal2013,roepkesim2018}
for additional discussion.

@@

Within this progenitor setting, many mechanisms for the explosion have 
been studied with the result being that the outcome is sensitive to
the ignition conditions and the nature of the thermonuclear buring.
Studies indicate that neither a pure deflagration (subsonic flame front)
nor a pure detonation (supersonic burning front) produce ejecta conisitent
with observations \citep{arnett69,roepkeetal07}.  Instead, the models 
that best reproduce the observed stratified ejecta are those in which 
the burning begins as a subsonic deflagration, which allows the star 
to react and expand thus lowering the density, and is then followed by a 
supersonic detonation that incinerates the 
star~\citep{Nomo84,Khokhlov1991Delayed-detonat,HoefKhok96,GameKhokOran05}.
The ``classic'' delayed detonation model is that of 
\citet{Khokhlov1991Delayed-detonat} \citep[See also][]{hoflich.khokhlov.ea:delayed,GameKhokOran05},
and many variations have been 
explored \citep[and references therein]{hillebrandtetal2013,calderetal2013}.
The approach we employ for this study is a variation, the 
deflagration-to-detonation transition (DDT), in which the transition
occurs at the top of a rising Rayleigh-Taylor unstable plume of hot, burned 
material when it reaches a threshold density.
We describe this approach and our methodology in detail below.

The double detionation picture is a variation of the single degenerate picture.
In this case, a white dwarf accretes material from a companion, but rather
than gaining enough mass to approach the Chandrasekhar limit, a detonation
occurs in the accreted layer that subsequently triggers a detonation
in the underlying white dwarf \citet{woosleyweavertaam80,taam80a,taam80b,
nomoto80,nomoto82b}. Early studies indicated that a detonation in an 
accreted He
layer could produce an inwardly propagating
shock that would ignite a detonation in the C-O core and found
that this scenario could work for a wide range of white dwarf
masses, not just the near-Chandrasekhar case \citep{livne90}, hence the
moniker ``sub-Chandrasekhar'' \citep{ww94}. Multidimensional supported the
efficacy of this picture~\citep{livneglasner91, livnearnett95,HoefKhok96,
hoeflichetal96, wigginsfalle97,wigginsetal98,garciasenzbravowoosley99}. 
These studies indicated that uncertainties in the initial conditions, 
however, were found to play a significant role on the explosion outcome.

A particular problem was that most models included a massive He layer on 
the white dwarf. Such a configuration leads to the unobserved result 
intermediate and heavy elements synthesized in the He detonation appear in 
the outer parts of the ejecta, which does not match obesrvation~\citep{HoefKhok96,
hoeflichetal96,finkhillebrandtroepke2007,simetal2010}.  \citet{bildstenetal2007}, 
however, found that fairly thin He layers could flash on sub-Chandrasekhar
mass white dwarfs, partially resolving the issue and encouraging
further research~\citep{simetal2012,brooksetal2015, shenetal2018,
glasneretal2018}.

The white dwarf merger progentor picture has two white dwarfs coming 
together and subsequently exploding~\citep{tutukovyungelson76,tutukovyungelson79,
webbink84,ibentutukov84}. This scenario provides an abundance of degenerate fuel, 
which may explain some bright events~\citep{scalzo:2010,Yuan:2010}. 
An early concern about this model came from studies finding that 
the as the stars merge, the more massive white dwarf can ignite near
its edge and a flame will burn inward, converting the C-O white dwarf into 
an O-Ne-Mg white dwarf~\citep{saionomoto1985,saionomoto2004}.  Further accretion 
leads to the collapse of the white dwarf
into a neutron star, a scenario termed ``accretion-induced collapse''
\citep{nomotokondo1991}.

Subsequent work allayed this concern~\citep{yoonetal2007,lorenaguilaretal2009,
Shenetal12, pakmoretal2012b} and motivated by the need to expain bright events
and the estimated population WD pairs, active research continues. Contemporary
research focuses on variations on the merger idea, including inspiraling pairs, 
collisions, violent mergers, and the ``core-degenerate'' model in which the merger 
takes place in a common envelope \citep{raskinetal2009,pakmoretal2011,kashi:2011,
pakmoretal2012a,Shenetal12,katzetal2016}.

@@

Our single-degenerate models are within the DDT explosion paradigm \cite{1986SvAL,
Khokhlov1991Delayed-detonat,NiemWoos97,Niem99,belletal2004,fishjump2015}. 
In this case, the accretion of mass on the white dwarf compresses and 
heats the core, igniting carbon fusion and driving a period of convection
\citep{WoosWunsKuhl04,wunschwoosley2004,Kuhletal06,nonakaetal2012}.
At some point, the fusion rate becomes fast enough due to the rising
temperature that energy production exceeds convective cooling and
the deflagration phase begins in the core~\citep{Nomo84,WoosWunsKuhl04}.
This flame is unstable, and as the deflagration propagates toward the surface
of the WD, it is subject to the Rayleigh-Taylor instability that generates
turbulence and boosts buring.  
Burning proceeds as a deflagration for about one second, and then 
the flame transitions to a detonation~\citep{hoflich.khokhlov.ea:delayed}.

@@

The mechanism by which the deflagration transistions to a detonation
is imcompletely understood, but the development of fluid instabilities
at the deflagration front, particuarly the Rayleigh-Taylor 
instability~\citep{taylor+50,chandra+81}, is an accepted part of the process.
Many ideas have been proposed, including the deflagration front
entering a regime of distributed burning and the net
rate becoming supersonic \citep{NiemWoos97}, the unstable
flame reaching a certian fractal dimension~\citep{woosley90},
and the Zel'dovich mechanism in which condtions are right for
a pocket of fuel to 
detonate~\citep{zeldovichetal1970,KhokOranWhee97,jacketal2014}.
We note that regardless of
the mechanism, the amount of expansion of the star prior to the bulk of burning
is critical to the composition of material synthesized in the explosion. Also,
early burning during the deflagtation phase is at high enough densites that
the effects of electron capture are significant and similarly influence the 
compositon the material synthesised in the explosion~\citep{hoeflichetal2004,
hoeflich2006,fesenetal2007,diamondetal2018}.
Our simulations assume that the transition occurs when the top of a 
rising, Rayleigh-Taylor unstable plume reaches a characteristic low
density~\citep{townsley.calder.ea:flame}. 
Variations on the choice of this density allowed earlier studies to explore
the effects of physics influencing the transition 
density~\citep{jacketal2010,Krueger2010On-Variations-o,kruegetal12}.

@@

Modern computing resources now enable simulations with unprecedented 
realism, allowing both one-dimensional simulations with a vast
amount of included physics and full three-dimensional simulations
albeit with less included physics~\citep{caldertownsley2018}.
In the area of stellar evolution, recent investigations revisiting 
late-time evolution of roughly 8 \Msun stars indicate that under
the right circumstances, ``hybrid" white dwarfs having a C/O core surrounded by O/Ne 
mantle may form \citep{siess2009,denissenkovetal2013}. These hybrid
white dwarf are thought to form when mixing at the lower convective boundary quenches
C burning in an asymptotic giant branch (AGB) star, leaving unburned C
in the core. The situation is at best uncertain, however, and the 
results depending on assumptions about convective
overshoot that have been questioned \citep{chenetal2014,lecoanetetal16,lattanzioetal2017}.

A hybrid WD can become a progenitor of a thermonuclear supernova via two 
of the progenitor system. 
First, the hybrid could be part of a binary system in which the companion star becomes
another white dwarf and the two merge. Recent population synthesis work supports
this pathway by indicating a substantial contribution to the Ia population rate from 
mergers where one member is a hybrid~\citep{yungelsonkuranov2017}. Second, if the hybrid 
WD is part of a binary with a main sequence or giant companion, it can gain mass and 
approach the Chandrasekhar mass, i.e.\  the single degenerate picture~\citep{willcoxetal2016}. 
{\color{red} COULD NOT FIND ANY EVIDENCE OF HYBRIDS IN THE SUBCHANDRA PATH. SAY 
SOMETHING?}

In either {\color{red}ANY} case, the hybrid WD will experience a period of cooling, which will
make the core-mantle interface subject to convective 
instability \citep{brooksetal2017,schwabgaraud2018}. In the case of the 
hybrid WD accreting and approaching the Chandrasekhar mass, accretion will
heat the core and start C fusion, leading to a period of ``simmering" prior
to the explosion \citep{PiroBild08}. The upshot is that is likely to be considerable
mixing after the hybrid forms that may homegenize the compostion 
\citep{denissenkovetal2015,brooksetal2017,schwabgaraud2018}.

@@

@@

The methodology of our study follows the approach of \citet{willcoxetal2016}. 
We per
Summarize what we do a la:
In our simulations, we initialize a detonation once the 
deflagration front reaches a
characteristic DDT fuel density, which controls the degree of
expansion the star undergoes during the deflagration stage.
